PineAPPL:\\ EW and mixed QCD--EW corrections for PDF fits

### 2020-08-01 paper/sec-results.tex
Typos

We focus on the following three processes: DY lepton-pair production, top-quark
pair production, and Z-boson (lepton-pair) production with non-zero transverse
momentum in proton-proton collisions. For each of these processes, we consider the measurements
described below.

\paragraph{DY lepton pair production.}
We select the distribution, single-differential in the invariant mass of the
lepton pair, $M_{\ell \bar\ell}$, measured by the ATLAS experiment at a centre-of
mass energy of 7~TeV in the high-mass region
($M_{\ell\bar\ell}>116$~GeV)~\cite{Aad:2013iua}.
We also select the distribution, double-differential in the rapidity and in
the invariant mass of the lepton pair, $y_{\ell\bar\ell}$ and $M_{\ell\bar\ell}$,
measured by the CMS experiment at a centre-of-mass energy of
7~TeV~\cite{Chatrchyan:2013tia}.
These measurements are currently included as standard in the
NNPDF3.1~\cite{Ball:2017nwa} and MMHT2014~\cite{Harland-Lang:2014zoa} PDF sets,
although with appropriate kinematic cuts that remove the bins at the largest
values of invariant mass, where EW corrections become sizeable. As explained in
Sect.~\ref{sec:pineappl-example}, should PI contributions be taken into account,
the process has a single LO, $\mathcal{O}(\alpha^2)$; at NLO, the
QCD contribution is $\mathcal{O}(\alpha_s\alpha^2)$, while the EW contribution
is $\mathcal{O}(\alpha^3)$. The latter is not included in the NLO
QCD computation, but it is in the NLO QCD+EW computation. Mixed QCD-EW
corrections occur only at NNLO, and are therefore not considered here.
EW corrections for this process were computed in
Refs.~\cite{Baur:2001ze,Dittmaier:2009cr}. The process receives contributions
from 12 (34) parton luminosities at NLO QCD (NLO QCD+EW).

@@

The central values of the renormalisation and factorisation scales, $\mu_R$ and
$\mu_F$, are chosen, for each process, as follows. In the case of DY lepton pair
production, we use the fixed scale $\mu_R=\mu_F=M_Z$, where $M_Z$ is the mass
of the $Z$-boson, for the ATLAS measurement, and the scale
$\mu_R=\mu_F=M_{\ell\bar\ell}$, where $M_{\ell\bar\ell}$ is the central value of each
invariant mass bin, for the CMS measurement.
In the case of top-quark pair production, we use the dynamic scales
$\mu_R=\mu_F=\sqrt{m_t^2+(p_T^t)^2}{\Big /}2$ for the distribution differential
in the transverse momentum of the top quark, and $\mu_R=\mu_F=H_T/4$ for the
distribution differential in the invariant mass of the top-quark pair, where
$H_T=\sqrt{m_t^2+(p_T^t)^2}+\sqrt{m_t^2+(p_T^{\bar{t}})}$, with $m_t$,
$p_T^t$ and $p_T^{\bar t}$ the mass of the top quark and the transverse momenta
of the top and antitop quarks, respectively. These choices were demonstrated
to maximise the convergence of the perturbative expansion~\cite{Czakon:2016dgf}.
In the case of $Z$-boson production with non-zero transverse momentum, we use
$\mu_R=\mu_F=M_Z$. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted in the Monte Carlo generation upon scale
variations. As is customary, the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and the envelope
from the nine-point scale variations is constructed.

@@

Finally, we implement the kinematic cuts specified in the corresponding
experimental analyses. In the case of high-mass DY lepton pair
production at 7~TeV measured by ATLAS, we require $p_T^\ell>25$~GeV,
$|\eta_\ell|<2.5$ and 116~GeV$<M_{\ell\bar\ell}<$ 1500~GeV for the transverse
momentum and the rapidity of each lepton and for the invariant mass of the
lepton pair, respectively. In the case of the DY lepton-pair production at
7~TeV measured by CMS, we require $p_T^{\ell_1}>14$~GeV, $p_T^{\ell_2}>9$~GeV,
$|\eta_\ell|<2.4$, $|\eta_{\ell\bar\ell}|<2.4$ and 20~GeV$<M_{\ell\bar\ell}<$ 1500~GeV
for the transverse momentum and the rapidity of each lepton, and for the
rapidity and the invariant mass of the lepton pair. In the case of $Z$-boson
production with non-zero transverse momentum at 13~TeV measured by CMS, we
require $p_T^\ell>25$~GeV, $|\eta_\ell|<2.4$,
$M_Z-15$~GeV$<M_{\ell\bar\ell}<M_Z+20$~GeV,
$|\eta_{\ell\bar\ell}|<2.4$ and 20~GeV$<p_T^{\ell\bar\ell}<1500$~GeV for the
transverse momentum and rapidity of each lepton, and for the invariant mass,
rapidity and transverse momentum of the lepton pair.

@@

\paragraph{DY lepton pair production.}

@@

%-------------------------------------------------------------------------------
\begin{figure}[!t]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLASZHIGHMASS49FB}
    \caption{Validation and test of the \textsc{PineAPPL} grid for the ATLAS
      high-mass DY lepton pair measurement in the high-mass region at
      a centre-of-mass energy of 7 TeV~\cite{Aad:2013iua}. The first panel
      displays the relative difference (in permille) between the {\sc PineAPPL}
      and the {\sc MC} results for the central, upper and lower scale choices,
      for theories accurate to both NLO QCD and NLO QCD+EW. The second, third
      and fourth panels present the theoretical predictions, accurate to either
      NLO QCD or NLO QCD+EW, always normalised to the former; on top of the
      theoretical predictions, the PDF uncertainty, the scale uncertainty and
      the Monte Carlo uncertainty are displayed in turn. The relative
      uncertainty of the experimental data is also shown for comparison.}
    \label{fig:atlaszhighmass49fb}
\end{figure}
%-------------------------------------------------------------------------------

@@

%-------------------------------------------------------------------------------
\begin{figure}[!p]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin3}%
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin4}\\
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin5}%
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin6}\\
    \caption{Same as Fig.~\ref{fig:atlaszhighmass49fb}, but for the CMS
      double-differential DY lepton pair measurement at a centre-of-mass energy
      of 7~TeV~\cite{Chatrchyan:2013tia}. Displayed are only two of the six
      invariant mass bins available, respectively below the $Z$-boson mass peak,
      45~GeV$<M_{\ell\bar\ell}<$60~GeV, on the $Z$-boson mass peak,
      60~GeV$<M_{\ell\bar\ell}<$120~GeV, above the mass peak,
      120~GeV$<M_{\ell\bar\ell}<$200~GeV, and at very high invarinat masses,
      200~GeV$<M_{\ell\bar\ell}<$1500~GeV.}
    \label{fig:cmsdy2d11_bins3456}
\end{figure}
%-------------------------------------------------------------------------------

As in the case of the ATLAS high-mass DY measurement, the CMS measurement is
also dominated by $q\bar{q}$ scattering. The leading (next-to-leading)
contribution to the 45~GeV$<M_{\ell\bar\ell}<$60~GeV invariant mass bin comes from
the $u\bar{u}$ ($d\bar{d}$) parton luminosity, which accounts for about 70\%
(22\%) of the double differential cross section, with small fluctuations across
the rapidity range. The PI contribution decreases from about 4\% at zero
rapidity to 1.5\% in the largest rapidity bin. The situation is slightly
different in the 60~GeV$<M_{\ell\bar\ell}<$120~GeV invariant mass bin, where the
leading (next-to-leading) contribution comes instead from the $d\bar{d}$
($u\bar{u}$) parton luminosity, which accounts for about 60\% (44\%) of the
double differential cross section at small rapidities, and for about 56\% (50\%)
at large rapidities. In the remaining two invariant mass bins, the leading
(next-to-leading) contribution comes again from the $u\bar{u}$ ($d\bar{d}$)
parton luminosity, which accounts for about 69\%-95\% (38\%-30\%) and
57\%-70\% (48\%-34\%), respectively for 120~GeV$<M_{\ell\bar\ell}<$200~GeV and
200~GeV$<M_{\ell\bar\ell}<$1500~GeV in the corresponding rapidity intervals; PI
contributions range between 7.3\%-1.2\% and 3.7\%-0.6\% in the two invariant
mass bins, respectively, for increasing rapidity.

>>>>

We focus on the following three processes: Drell--Yan lepton-pair production, top-quark
pair production, and Z-boson (lepton-pair) production with non-zero transverse
momentum in proton-proton collisions. For each of these processes, we consider the measurements
described below.

\paragraph{Drell--Yan lepton pair production.}
We select the distribution, single-differential in the invariant mass of the
lepton pair, $M_{\ell \bar\ell}$, measured by the ATLAS experiment at a centre-of
mass energy of 7~TeV in the high-mass region
($M_{\ell\bar\ell}>116$~GeV)~\cite{Aad:2013iua}.
We also select the distribution, double-differential in the rapidity and in
the invariant mass of the lepton pair, $y_{\ell\bar\ell}$ and $M_{\ell\bar\ell}$,
measured by the CMS experiment at a centre-of-mass energy of
7~TeV~\cite{Chatrchyan:2013tia}.
These measurements are currently included as standard in the
NNPDF3.1~\cite{Ball:2017nwa} and MMHT2014~\cite{Harland-Lang:2014zoa} PDF sets,
although with appropriate kinematic cuts that remove the bins at the largest
values of invariant mass, where EW corrections become sizeable. As explained in
Sect.~\ref{sec:pineappl-example}, should PI contributions be taken into account,
the process has a single LO, $\mathcal{O}(\alpha^2)$; at NLO, the
QCD contribution is $\mathcal{O}(\alpha_s\alpha^2)$, while the EW contribution
is $\mathcal{O}(\alpha^3)$. The latter is not included in the NLO
QCD computation, but it is in the NLO QCD+EW computation. Mixed QCD-EW
corrections occur only at NNLO, and are therefore not considered here.
EW corrections for this process were computed in
Refs.~\cite{Baur:2001ze,Dittmaier:2009cr}. The process receives contributions
from 12 (34) parton luminosities at NLO QCD (NLO QCD+EW).

@@

The central values of the renormalisation and factorisation scales, $\mu_R$ and
$\mu_F$, are chosen, for each process, as follows. In the case of Drell--Yan lepton pair
production, we use the fixed scale $\mu_R=\mu_F=M_Z$, where $M_Z$ is the mass
of the $Z$-boson, for the ATLAS measurement, and the scale
$\mu_R=\mu_F=M_{\ell\bar\ell}$, where $M_{\ell\bar\ell}$ is the central value of each
invariant mass bin, for the CMS measurement.
In the case of top-quark pair production, we use the dynamic scales
$\mu_R=\mu_F=\sqrt{m_t^2+(p_T^t)^2}{\Big /}2$ for the distribution differential
in the transverse momentum of the top quark, and $\mu_R=\mu_F=H_T/4$ for the
distribution differential in the invariant mass of the top-quark pair, where
$H_T=\sqrt{m_t^2+(p_T^t)^2}+\sqrt{m_t^2+(p_T^{\bar{t}})}$, with $m_t$,
$p_T^t$ and $p_T^{\bar t}$ the mass of the top quark and the transverse momenta
of the top and antitop quarks, respectively. These choices were demonstrated
to maximise the convergence of the perturbative expansion~\cite{Czakon:2016dgf}.
In the case of $Z$-boson production with non-zero transverse momentum, we use
$\mu_R=\mu_F=M_Z$. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted in the Monte Carlo generation upon scale
variations. As is customary, the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and the envelope
from the nine-point scale variations is constructed.

@@

Finally, we implement the kinematic cuts specified in the corresponding
experimental analyses. In the case of high-mass Drell--Yan lepton pair
production at 7~TeV measured by ATLAS, we require $p_T^\ell>25$~GeV,
$|\eta_\ell|<2.5$ and 116~GeV$<M_{\ell\bar\ell}<$ 1500~GeV for the transverse
momentum and the rapidity of each lepton and for the invariant mass of the
lepton pair, respectively. In the case of the Drell--Yan lepton-pair production at
7~TeV measured by CMS, we require $p_T^{\ell_1}>14$~GeV, $p_T^{\ell_2}>9$~GeV,
$|\eta_\ell|<2.4$, $|\eta_{\ell\bar\ell}|<2.4$ and 20~GeV$<M_{\ell\bar\ell}<$ 1500~GeV
for the transverse momentum and the rapidity of each lepton, and for the
rapidity and the invariant mass of the lepton pair. In the case of $Z$-boson
production with non-zero transverse momentum at 13~TeV measured by CMS, we
require $p_T^\ell>25$~GeV, $|\eta_\ell|<2.4$,
$M_Z-15$~GeV$<M_{\ell\bar\ell}<M_Z+20$~GeV,
$|\eta_{\ell\bar\ell}|<2.4$ and 20~GeV$<p_T^{\ell\bar\ell}<1500$~GeV for the
transverse momentum and rapidity of each lepton, and for the invariant mass,
rapidity and transverse momentum of the lepton pair.

@@

\paragraph{Drell--Yan lepton pair production.}

@@

%-------------------------------------------------------------------------------
\begin{figure}[!t]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLASZHIGHMASS49FB}
    \caption{Validation and test of the \textsc{PineAPPL} grid for the ATLAS
      high-mass Drell--Yan lepton pair measurement in the high-mass region at
      a centre-of-mass energy of 7 TeV~\cite{Aad:2013iua}. The first panel
      displays the relative difference (in permille) between the {\sc PineAPPL}
      and the {\sc MC} results for the central, upper and lower scale choices,
      for theories accurate to both NLO QCD and NLO QCD+EW. The second, third
      and fourth panels present the theoretical predictions, accurate to either
      NLO QCD or NLO QCD+EW, always normalised to the former; on top of the
      theoretical predictions, the PDF uncertainty, the scale uncertainty and
      the Monte Carlo uncertainty are displayed in turn. The relative
      uncertainty of the experimental data is also shown for comparison.}
    \label{fig:atlaszhighmass49fb}
\end{figure}
%-------------------------------------------------------------------------------

@@

%-------------------------------------------------------------------------------
\begin{figure}[!p]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin3}%
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin4}\\
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin5}%
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin6}\\
    \caption{Same as Fig.~\ref{fig:atlaszhighmass49fb}, but for the CMS
      double-differential Drell--Yan lepton pair measurement at a centre-of-mass energy
      of 7~TeV~\cite{Chatrchyan:2013tia}. Displayed are only two of the six
      invariant mass bins available, respectively below the $Z$-boson mass peak,
      45~GeV$<M_{\ell\bar\ell}<$60~GeV, on the $Z$-boson mass peak,
      60~GeV$<M_{\ell\bar\ell}<$120~GeV, above the mass peak,
      120~GeV$<M_{\ell\bar\ell}<$200~GeV, and at very high invarinat masses,
      200~GeV$<M_{\ell\bar\ell}<$1500~GeV.}
    \label{fig:cmsdy2d11_bins3456}
\end{figure}
%-------------------------------------------------------------------------------

As in the case of the ATLAS high-mass Drell--Yan measurement, the CMS measurement is
also dominated by $q\bar{q}$ scattering. The leading (next-to-leading)
contribution to the 45~GeV$<M_{\ell\bar\ell}<$60~GeV invariant mass bin comes from
the $u\bar{u}$ ($d\bar{d}$) parton luminosity, which accounts for about 70\%
(22\%) of the double differential cross section, with small fluctuations across
the rapidity range. The PI contribution decreases from about 4\% at zero
rapidity to 1.5\% in the largest rapidity bin. The situation is slightly
different in the 60~GeV$<M_{\ell\bar\ell}<$120~GeV invariant mass bin, where the
leading (next-to-leading) contribution comes instead from the $d\bar{d}$
($u\bar{u}$) parton luminosity, which accounts for about 60\% (44\%) of the
double differential cross section at small rapidities, and for about 56\% (50\%)
at large rapidities. In the remaining two invariant mass bins, the leading
(next-to-leading) contribution comes again from the $u\bar{u}$ ($d\bar{d}$)
parton luminosity, which accounts for about 69\%-95\% (38\%-30\%) and
57\%-70\% (48\%-34\%), respectively for 120~GeV$<M_{\ell\bar\ell}<$200~GeV and
200~GeV$<M_{\ell\bar\ell}<$1500~GeV in the corresponding rapidity intervals; PI
contributions range between 7.3\%-1.2\% and 3.7\%-0.6\% in the two invariant
mass bins, respectively, for increasing rapidity.
